Foremost,  I would like to thank my committee members: Dr.~Jack
Snoeyink, Dr.~Montek Singh, Dr.~David Evans, Dr.~Ilya Mironov for
agreeing to trudge through my dissertation. I have my deepest
appreciation to my advisor, Dr.~Michael Reiter, for providing me
enough support in my research. His endless enthusiasm and rigorous
attitude always help me make our work better than expected. Our weekly
meetings would be the most valuable moments.  Being Mike's student is
undoubtedly the correct choice I made.

I also want to acknowledge my collaborators and mentors. Although I am
the only author of this dissertation, I get rid of ``I" in chapters to
credit my advisor and my collaborators' contributions. I thank
Dr.~Yinqian Zhang for sharing his research experience to help me get
started and providing useful feedback about the system design, when
he was a postdoctoral researcher in the same group and even after
joining OSU. I thank to Dr.~Zhiyun Qian for his active collaboration
in static analysis and his insightful feedback in my case studies.
Many thanks to our Intel contacts, Dr.~Matthew Fernandez for his
encourage in my research and his voluntary help in my job searching.
Special thanks to my mentor Dr.~Junghwan Rhee for his patient
supervision in NEC Lab.  The intern experience with him provides me
enough chance to learn from machine learning experts, which highly
broadens my research horizon. I appreciate my mentor, Dr.~Michael
Vrable, at Google, for providing me career advice and support. With
his help, I learned the value of industry problems, which are easy for
small systems but become tricky in large systems.

I thank my labmates and my friends at UNC-Chapel Hill. With their
help, I do not feel that I am alone. In particular, thanks to Andrew Chi,
Robert Cochran, Adam Humphries, Victor Heorhiadi, Jung Jiang,  Sheng
Liu, Marie Nesfield, Ke Coby Wang, and Qiuyu Xiao. I always learned a
lot from their works from different perspectives. 

A very special gratitude goes out to National Science Foundation and
Intel for funding the work.

Completing my Ph.D. while leaving my hometown is not easy for me. The
key making me go so far with comfort is that I have my family,
particularly my father, my wonderful stepmother, and my sister support
and encourage me. I owe a lot to my grandmother, who raised me during
my childhood. I thank Sheng Liu, who knows and supports me much no
matter what happens. I could never have done this without them. I love
you all.
